---
title: "Answers to Socrative Quiz - Lesson 3"
author: "EA"
date: "January 25, 2019"
output: html_document
---
<br/>

1. Which latrine depth has the greatest mean number of OTUs?
  
```{r eval=FALSE}
split_dat %>% 
  group_by(Depth) %>% 
  summarize(mean = mean(OTUs)) %>% 
  arrange(desc(mean))
```
__9 cm__ is the latrine depth with the greatest mean number of OTUs.

<br/>

2. Is there more Clostridia on average in sites from Tanzania or Vietnam?


```{r eval = FALSE}
split_dat %>%
  filter(Taxa == "Clostridia") %>%
  group_by(Country) %>%
  summarise(mean = mean(OTUs)) %>%
  arrange(desc(mean))
```

There is more Clostridia in __Tanzania__.

<br/>

3. Which site had the greatest number of Taxa represented?


```{r eval = FALSE}
split_dat %>% 
  group_by(Country, Latrine_Number, Depth) %>% 
  filter(OTUs != 0) %>%
  summarize(count = n()) %>% 
  arrange(desc(count))
```


__V_7_3__ has the greatest number of Taxa represented.

<br/>

4. The gapminder_wide dataset ignores the following rules of tidy data:

__a__. Multiple variables are in one column      
__&__ 
__c__. Variables are stored in both rows and columns.

<br/>

5. How many rows will the final gapminder data frame have?

```{r eval = FALSE}
gap %>% 
  gather(obs_type, obs_value, -continent, -country) %>% 
  separate(obs_type, c("obs_type", "year"), sep="_", convert = TRUE) %>% 
  spread(key = obs_type, value = obs_value) %>% 
  select(continent, country, year, lifeExp, pop, gdpPercap) %>%
  nrow()
```

The final gapminder data frame will have __1704__ rows.

<br/>

6. What is/are the resulting data type(s) of this vector?

c(Samples = "V_2_10", Temp = 5, pH = 6, TS = 5)

<br/>

Each element in the vector will be coerced to __character__, which is the most inclusive data type out of integer and character.

<br/>

7. Make a column and add it to jdat using the dplyr function bind_cols().

```{r eval = FALSE}
#one possible answer 
jdat <- bind_cols(jdat, new_col = rep(1, nrow(jdat)))
```

<br/>

8. What is the total number of characters in the trilogy?

```{r eval=FALSE}
#read in the book data frames, which contain the character info
fellowship <- read_csv("data/fellowship.csv")
towers <- read_csv("data/towers.csv")
return <- read_csv("data/returnking.csv")

#bind data frames together to get characters from all 3 books
lotr <- fellowship %>% bind_rows(towers, return) 

#get the unique set of characters
lotr %>% select(Character) %>% unique() %>% nrow()
```
There are __68__ characters in the trilogy.

<br/>

9. Which race speaks the most?

```{r eval=FALSE}
#gather Races into tidy data format
lotr <- lotr %>% gather(key = Race, value = Words, Dwarf:Wizard, na.rm=TRUE) 

lotr %>% 
  group_by(Race) %>% 
  summarize(sum = sum(Words)) %>% 
  top_n(1) #or arrange(desc(sum))
```

__Hobbits__ speak the most overall.

<br/>

10. True or false. Hobbits speak the most in all 3 films.

```{r eval = FALSE}
lotr %>% 
  group_by(Film, Race) %>% 
  summarize(sum = sum(Words)) %>% 
  top_n(1)
```

__False__. Men speak more than hobbits in 2 of the 3 films.

<br/>

11. Which character is the chattiest?


```{r eval = FALSE}
lotr %>% 
  group_by(Character) %>% 
  summarize(sum = sum(Words)) %>% 
  top_n(1)
```
__Gandalf__ is the character that speaks the most.

<br/>

12. Which gender speaks the most for Elvish characters?

```{r eval = FALSE}
#read in the gender datasets
female <- read_csv("data/Female.csv")
male <- read_csv("data/Male.csv")
#bind the datasets together
gender <- bind_rows(female, male) %>%
#gather Races together
  gather(Race, Words, Elf:Man, na.rm=T)

gender %>% 
  filter(Race == "Elf") %>% 
  group_by(Gender) %>% 
  summarize(sum = sum(Words)) %>%
  arrange(desc(sum))
```

Overall, __male__ elves speak more than female elves.

<br/>

13. True or false. Male elves speak more than female elves across all 3 films.

```{r eval=FALSE}
gender %>% 
  filter(Race == "Elf") %>% 
  group_by(Film, Gender) %>% 
  summarize(sum = sum(Words)) %>%
  top_n(1)
```
__False__. Female elves speak more than male elves in one of the three films.

<br/>

14. Who is the top speaking elf in each film?

```{r eval=FALSE}
lotr %>% 
  filter(Race == "Elf") %>% 
  group_by(Film, Character) %>% 
  summarize(sum = sum(Words)) %>% 
  top_n(1)
```
__Galadriel, Erlond, Legolas__ are the top speaking elves in The Fellowship of the Ring, The Two Towers, and The Return of the King, respectively.



